---
title: "Color Palettes in stemtool"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(stemtools)
library(scales)
```

`stemtool` provides a set of color palettes for plotting. There are three types of palettes:

-   **Diverging** - used for variables where low values are the opposite of high values. Typically likert items, where low values signify agreement and high values signify disagreement.

    -   Available diverging palettes: `modern`, `gruvbox_div`.

-   **Nominal** - used for variables where categories have no set order. Typically gender, countries of origin or occupation.

    -   Available nominal palettes: `gruvbox`.

-   **Sequential** - used for variables where higher values indicate higher frequency/concentration of something. Typically level of unemployment, number of people with tertiary education or socioeconomic class.

    -   Available sequential palettes: none (for now).

## Accesing `stemtool` palettes

Color codes from palettes can accessed using `stem_palettes()` function. The function expects palette name (see above) and optionally the number of colors to extract. By default, all colors are extracted. Palettes can be visually inspected with `show_col()` function from `scales` package:

```{r stem-palettes-codes}
stem_palettes(palette = "modern", n = 5)
show_col(stem_palettes(palette = "modern"))
```

## Using palettes in `ggplot2`

Palettes can be used in `ggplot2` figures by using functions `scale_color_stem()` and `scale_fill_stem()`. Specific palette can be chosen by `palette` argument, using `direction = -1` can be used to reverse the order of colors. Note that for diverging palettes, colors are used from both ends of the palette in an alternating pattern.

```{r ggplot-example}
mtcars |> 
  within(gear <- as.factor(gear)) |> 
  ggplot(aes(x = mpg,
             y = hp,
             color = gear)) +
  geom_point(size = 3) +
  scale_color_stem(palette = "modern")
```
