---
title: "Colour palettes and the Stem theme"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Colour palettes and the Stem theme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 150
)
```

```{r setup, message=FALSE, warning=FALSE}
library(stemtools)
library(ggplot2)
library(scales)
```

`{stemtools}` ships a set of colour palettes and a complete `{ggplot2}` theme that together give plots the Stem look. This vignette shows how to browse the palettes, use them the same way you would any other palette in `{ggplot2}`, and how to switch on `theme_stem()`.

## Available palettes

Every palette lives in a single registry, and each one is tagged with a *type* that all the other palette functions read from. There are three types:

-   **Nominal** -- for variables whose categories have no inherent order (gender, country of origin, occupation).

    -   Available nominal palettes: `nom1`, `nom2`.

-   **Sequential** -- for ordered variables where higher values mean *more* of something (level of unemployment, share with tertiary education, socioeconomic class).

    -   Available sequential palettes: `seq1`, `seq2`, `seq3`, `seq4`.

-   **Diverging** -- for variables with a meaningful midpoint, where low values are the opposite of high values (typically Likert items running from agreement to disagreement).

    -   Available diverging palettes: `modern`, `div1`, `div2`, `div3`.

The quickest way to see everything at a glance is `stem_palettes_all()`, which draws every palette as a row of colour tiles, grouped by type (much like `RColorBrewer::display.brewer.all()`):

```{r stem-palettes-all, fig.height = 5}
stem_palettes_all()
```

## Accessing palette colours

The hex codes for a single palette are returned by `stem_palette()`, which takes a palette name and returns all of its colours. The returned vector carries a `type` attribute, and can be inspected visually with `show_col()` from `{scales}`:

```{r stem-palettes-codes}
stem_palette("modern")

attr(stem_palette("modern"), "type")

show_col(stem_palette("modern"))
```

To pull a specific number of colours, use the generator returned by `stem_palette_gen()`. For diverging palettes the colours are sampled symmetrically around the midpoint rather than simply taken from the front of the palette:

```{r stem-palette-gen}
stem_palette_gen("modern")(3)
```

## Using palettes in `{ggplot2}`

Stem palettes plug into `{ggplot2}` exactly like the built-in `scale_*_brewer()` or `scale_*_viridis_d()` scales. Reach for `scale_colour_stem()` (or its American alias `scale_color_stem()`) for the `colour` aesthetic and `scale_fill_stem()` for `fill`. Pick a palette with the `palette` argument and, if needed, flip its order with `direction = -1`:

```{r ggplot-colour}
mtcars |>
  within(gear <- as.factor(gear)) |>
  ggplot(aes(x = mpg, y = hp, colour = gear)) +
  geom_point(size = 3) +
  scale_colour_stem(palette = "modern")
```

Because these are ordinary discrete scales, any argument accepted by `ggplot2::discrete_scale()` -- `name`, `breaks`, `labels`, and so on -- can be passed straight through. Here is a `fill` scale using a nominal palette:

```{r ggplot-fill}
ggplot(mpg, aes(x = drv, fill = drv)) +
  geom_bar() +
  scale_fill_stem(palette = "nom2") +
  guides(fill = "none")
```

## The Stem theme

`theme_stem()` is a *complete* `{ggplot2}` theme carrying the Stem look: no gridlines, a top legend, bold titles and the Stem house font. Because it is complete, you activate it once with `ggplot2::theme_set()` and every subsequent plot picks it up:

```{r theme-set}
theme_set(theme_stem(family = ""))
```

Here `family = ""` falls back to the graphics device's default font, which is handy on machines where the Stem house font (Calibri) is not installed. In day-to-day use you can simply call `theme_stem()` with no arguments.

With the theme active, the same plot as above now carries the Stem styling automatically:

```{r theme-applied}
mtcars |>
  within(gear <- as.factor(gear)) |>
  ggplot(aes(x = mpg, y = hp, colour = gear)) +
  geom_point(size = 3) +
  labs(title = "Horsepower against fuel economy") +
  scale_colour_stem(palette = "modern")
```

You can also add the theme to a single plot with `+ theme_stem()`, and override individual settings through its arguments -- `ink` (foreground), `paper` (background), and `accent` -- or via `...`, which is forwarded to `ggplot2::theme()`:

```{r theme-custom}
ggplot(mpg, aes(x = drv, fill = drv)) +
  geom_bar() +
  scale_fill_stem(palette = "nom2") +
  labs(title = "Vehicles by drivetrain") +
  theme_stem(family = "", legend.position = "none")
```

Palettes and the theme are designed to be used together: the `stem_*` plotting functions rely on the same palettes, and pair naturally with `theme_stem()` set for the whole session.
