---
title: "Plotting with stemtools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting with stemtools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  dpi = 150
)
```

This vignette showcases every plotting function in `{stemtools}`: `stem_barplot()`, `stem_barstack()`, `stem_inline()`, `stem_battery()` and `stem_multiselect()`. We'll use the simulated `trust` dataset that ships with the package.

```{r setup}
library(stemtools)
library(ggplot2)

data(trust, package = "stemtools")
head(trust)
```

## Setting the Stem theme

`theme_stem()` is a *complete* theme carrying the Stem look. Because it is complete, the plotting functions do not apply it themselves -- instead we switch it on once with `theme_set()` and every plot from here on picks it up automatically:

```{r theme}
theme_set(theme_stem(family = ""))
```

We pass `family = ""` so the plots fall back to the graphics device's default font (the Stem house font, Calibri, may not be installed on every machine). In normal use you would just call `theme_set(theme_stem())`.

## How the plotting functions are built

All the plotting functions are composed from a shared frequency core. Each one calls `stem_summarise_cat()` to compute (possibly weighted) proportions with 95% confidence intervals, then hands the aggregated data to a `{ggplot2}` template. You can call the frequency function directly whenever you need the numbers themselves:

```{r summarise}
stem_summarise_cat(trust, item = police, group = eu_index, weight = W)
```

Every plotting function accepts the same core arguments:

* `item` -- the categorical variable to plot,
* `group` -- an optional grouping variable,
* `weight` -- optional survey weights (passed to `{surveycore}`),
* `palette` and `direction` -- the [Stem colour palette][stem_palette] to use,
* `label_*` -- controls for the percentage labels.

## Simple bar plot

`stem_barplot()` draws the distribution of a single categorical variable as horizontal bars, one bar per category.

```{r barplot}
stem_barplot(trust, item = government)
```

Supplying a `group` variable draws dodged bars so categories can be compared across groups. Proportions are then computed *within* each group.

```{r barplot-group}
stem_barplot(trust, item = police, group = eu_index, weight = W)
```

Confidence intervals are always computed and can be shown with `errorbar = TRUE`.

```{r barplot-errorbar}
stem_barplot(trust, item = government, errorbar = TRUE)
```

## Stacked bar plot

`stem_barstack()` draws one stacked horizontal bar per category of the `group` variable. Each bar sums to 100%, which makes it easy to compare the composition of the item across groups.

```{r barstack}
stem_barstack(trust, item = police, group = eu_index, weight = W)
```

## Inline bar plot

`stem_inline()` is a compact summary of a single variable as one stacked horizontal bar.

```{r inline, fig.height = 2}
stem_inline(trust, item = police, weight = W)
```

## Battery of like items

`stem_battery()` plots several variables that share the same response categories (a Likert battery, say) as one chart, with a stacked bar per item. Pass the items with tidyselect helpers and, optionally, `order_by` to sort the items by their combined share of one or more categories.

```{r battery}
stem_battery(trust,
             items = c(police, eu, government, army, scientists),
             weight = W,
             order_by = c("Definitely Agree", "Rather Agree"))
```

Item labels (the `"label"` attribute) are used automatically when present; set `item_label = FALSE` to use the bare variable names.

## Multiple-choice items

`stem_multiselect()` summarises a set of "select all that apply" items, rescaling the frequencies so the bars show the share of *respondents* who picked each option (rather than the share of responses).

```{r multiselect}
stem_multiselect(trust, items = dplyr::starts_with("biggest_concern"), weight = W)
```

Supplying a `group` variable draws dodged bars for a within-group comparison.

## Customizing labels

By default the plots print a percentage label on each bar (set `labels = FALSE` to hide them). `label_accuracy` controls rounding (`1` for whole numbers, `0.1` for one decimal place), `label_suffix` sets the text appended to each label and `label_hide` suppresses labels for proportions below a threshold.

```{r labels}
stem_barstack(trust,
              item = police,
              group = eu_index,
              label_accuracy = 0.1,
              label_hide = 0.1)
```

## Colours

Use `palette` to pick one of the [Stem palettes][stem_palette] and `direction = -1` to reverse it.

```{r palette}
stem_barstack(trust, item = police, group = eu_index, palette = "div2")
```

## Collapsing categories

`collapse_item` and `collapse_group` collapse (or rename) categories by passing a named list, exactly as in `stem_summarise_cat()`.

```{r collapse}
stem_barstack(trust,
              item = police,
              group = eu_index,
              collapse_item = list(`Ano` = c("Definitely Agree", "Rather Agree"),
                                   `Ani ano, ani ne` = "Neither Agree nor Disagree",
                                   `Ne` = c("Rather Disagree", "Definitely Disagree")))
```

## Weights

If a `weight` variable is specified, weights are treated as survey weights and passed into `{surveycore}`, which builds a Taylor-series survey design. Manipulating that design uses `{surveytidy}`, whose verbs mirror `{dplyr}`, so weighted and unweighted code read the same way.

With unweighted data, the confidence intervals are computed using the basic $\sqrt{\frac{p \cdot (1-p)}{n}}$ formula. With very small proportions and small samples this can lead to bounds outside of the `(0, 1)` range. With weighted data, the confidence intervals come from `{surveycore}` using Taylor-series linearization.

```{r weights}
stem_barplot(trust, item = government, weight = W)
```

## Further customization

Every function returns a standard `{ggplot2}` object, so you can keep adding layers, scales and theme tweaks. The global `theme_stem()` is already applied, so this just layers a title on top:

```{r customize}
stem_barplot(trust, item = government) +
  labs(title = "Trust in the government")
```

[stem_palette]: ../reference/stem_palette.html
