---
title: "Plotting with stemtools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{plotting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to use `{stemtools}` for creating a variety of standardized plots. We'll use a simuated dataset `trust`, included with the package itself.

```{r setup}
library(stemtools)

data(trust, package = "stemtools")
head(trust)
```

The core function is `stem_plot()`, a convinient wrapper around `stem_summarise()`, followed by a flexible `{ggplot2}` template. The aim of the `stem_plot()` is to visualize *aggregated* data, namely counts and means.

```{r stem-summarise-show}
stem_plot(data = trust, item = police)
```

`stem_plot()` automatically recognizes whether the item variable is numerical or categorical. For numerical variables, grouping variable should also be included using `group` argument. Survey weights can be used by calling the `weight` argument.

!TODO: FIX NUMERICAL PLOTS

## Customizing labels

By default, `stem_plot()` shows labels for each geom (set `labels = FALSE` to hide them). Looks of the labels can be customized using several arguments. `label_scale` multiples labels by specified amount (e.g. `100`). `label_accuracy` rounds labels - `1` for no decimal places, `0.1` for one decimal place, `0.01` for two, and so on. `label_suffix` and `label_prefix` allow to add suffix and prefix respectively. `label_big` controls thousands separator and `label_decimal` control which character is used to separate decimal digits. You can also hide labels lower than a certain threshold with `label_hide` (with `0.05` being the default). Additionally, you can pass a list into `label_args` to specify additional arguments like `vjust` or `position`:

```{r labels-example}
stem_plot(data = trust,
          item = government,
          label_scale = 100,
          label_suffix = "%",
          label_hide = 0)
```

## Customizing title, caption and axes

By default, plots include title. If the item has an attribute `label` (such as one created using the `{labelled}` package) and `title_label = TRUE`, the label will be used instead of variable name. You can use `title_wrap` to set how many characters before the title is wrapped on the next line. If you don't want to use title, set `title = FALSE`. If you want to use your own title, use `stem_plot(..., title = FALSE) + ggplot::labs(title = "My title")`.

By default, the plot also includes a caption with sample size. To get rid off the caption, set `caption = FALSE`.

```{r title-example}
stem_plot(data = trust,
          item = government,
          title_label = TRUE,
          caption = TRUE)
```

You can pass `{ggplot2}`'s scale functions into arguments `scale_x`, `scale_y`, `scale_fill` and `scale_color`. Note that you may need to set limits by hand when using grouping variable and setting your own scale. Similarly, you can control legend using `guides`.

Lastly, you can reverse the order of item and group variable using `item_reverse` and `group_reverse`.

```{r scales-example}
stem_plot(data = trust,
          item = police,
          group = eu_index,
          label = FALSE,
          scale_y = ggplot2::scale_y_continuous(labels = scales::percent_format(),
                                                limits = c(0, NA)),
          scale_fill = ggplot2::scale_fill_brewer(palette = "Set1"),
          scale_color = ggplot2::scale_color_brewer(palette = "Set1"),
          guides = ggplot2::guides(fill = ggplot2::guide_legend(reverse = TRUE),
                                   color = ggplot2::guide_legend(reverse = TRUE)),
          item_reverse = TRUE)
```

## Collapsing categories

You can use `collapse_item` and `collapse_group` to collapse variable categories by passing a named list. The same arguments can also be used to rename categories.

```{r}
stem_plot(data = trust,
          item = police,
          group = eu_index,
          label = FALSE,
          collapse_item = list(`Ano` = c("Definitely Agree",
                                         "Rather Agree"),
                               `Ani ano, ani ne` = "Neither Agree nor Disagree",
                               `Ne` = c("Rather Disagree",
                                        "Definitely Disagree")),
          collapse_group = list(`Pozitivní vztah` = "Likes EU",
                                `Neutrální vztah` = "Neutral",
                                `Negativní vztah` = "Dislikes EU",
                                `Neví` = "Doesn't Know"))
```

## Weights

If weights variable is specified, weights are treated as survey weights passed into `{svryr}` (which in turn passes it into `{survey}`).

```{r stem-plot-weight}
stem_plot(data = trust,
          item = government,
          label_scale = 100,
          weight = W)
```

## Choosing geoms

You can choose other geoms to represent the data. For all estimates, 95% confidence intervals are also computed and can be plotted using an appropriate geom. You can also set additional arguments for your geom by passing a named list into `geom_args`.

Note that with unweighted data, the confidence intervals are computed using the basic $\sqrt{\frac{p \cdot (1-p)}{n}}$ formula. With very small proportions and small sample sizes, this can lead to the confidence intervals' bounds being outside of realistic values (i.e. proportions lower than 0 or higher than 1). If the data are weighted, the confidence intervals are based on logistic regression and will always be properly bounded.

```{r geom-example}
stem_plot(data = trust,
          item = police,
          group = eu_index,
          weight = W,
          label = FALSE,
          geom = ggplot2::geom_pointrange)
```

## 
